---
output:
  word_document: default
  html_document: default
---

## Name :- Aggarwal Ravindra 
## Project 1 
## Course - BAN 502 

Loading Libraries 

```{r}
options(tidyverse.quiet=TRUE)
#install.packages("dplyr")
library(dplyr)
#install.packages("tidyverse")
library(tidyverse)
library(ggplot2)
library(GGally)
library(cluster) #algorithms for clustering
library(factoextra) #visualization
library(caret) #for the dummyVars function
library(mice) #package for imputation
#install.packages("VIM")
library(VIM) #visualizing missingness
library(randomForest) #random forests, using rather than ranger, bug as of Feb 2019 in ranger parameter tuning
library(e1071) #often needed for model building assistance
library(nnet) #for neural networks
library(rpart) #for classification trees
library(RColorBrewer) #better visualization of classification
library(leaps) #best subset selection
library(lubridate)
library(MASS)
library(rattle) #better visualization of classification trees
#install.packages("PerformanceAnalytics")
library(PerformanceAnalytics)
```

Reading Chicago 

```{r}
Chicago = read_csv ("Chicago.csv")
```

## Review and Prepare data

* Chicago data frame has 22 variables and 267185 observations of data. 

We are trying to predict " Arrest". This data frame has information of crime, location, police ID, person ID, type, date and location of crime, arrest made or not. 

Response Variable is "Arrest". 

We are deleting ID,Case Number, Updated On, X Coordinate, Y Coordinate amd Location. 


```{r}
str(Chicago)
summary(Chicago)
```
```{r}

Chicago = Chicago %>% dplyr:: select (-ID,-'Case Number',-'Updated On',-'X Coordinate',-'Y Coordinate', -'Location')

```


Check structure and summary again for missing values. 

```{r}
str(Chicago)
summary(Chicago)
```

After removing columns, there are 16 variables left in data set. Following variables have missing value 

" Ward" = 4, "Community Area" = 2, Latitude and Longitude have 4365 observations missing. 

All incident occured in 2018.

```{r}
vim_plot = aggr(Chicago, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
#the cex.axis reduces size of text on x-axis so labels fit better
```
```{r}
Chicago_NA <- Chicago %>% drop_na('Description',Ward,'Community Area',Latitude, Longitude)
```

```{r}
Chicago_NA$Description[Chicago_NA$Description == ""] = NA
Chicago_NA <- Chicago_NA %>% drop_na('Description')
```

Let us display again

```{r}
vim_plot1 = aggr(Chicago_NA, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
```


```{r}
str(Chicago_NA)
summary(Chicago_NA)
```

Find time from Dates


```{r}
Chicago_NA = Chicago_NA %>% mutate(Date=mdy_hms(Date))
```

Find Hours, and create new variable called Hour 

```{r}
Chicago_NA = Chicago_NA %>% mutate(Hour = hour(Date))# Creates new variable in dataset with the # hour extracted from each date/time
summary(Chicago_NA$Hour)

```


Extracting minutes from date, and adding new variable days.

```{r}
Chicago_NA = Chicago_NA %>% mutate(Minute = minute(Date))# Creates new variable in dataset with the # minute extracted from each date/time
summary(Chicago_NA$Minute)
```

All crime happened between 18.00 to 18.23 minutes in this database. 

Extracting Month from Date and adding new variable Month 
```{r}
Chicago_NA = Chicago_NA %>% mutate(Month = months(Date))# Creates new variable in dataset with the # Month extracted from each date/time
summary(Chicago_NA$Month)

```
Extracting Days from Date and adding new variable for Day 

```{r}
Chicago_NA = Chicago_NA %>% mutate(Day= day(Date))# Creates new variable in dataset with the # day extracted from each date/time
summary(Chicago_NA$Day)
```

Converting variables to factors. 

```{r}
Chicago_NA = Chicago_NA %>% mutate(Hour = as.factor(Hour))%>%mutate(`Primary Type`=as.factor(`Primary Type`))%>% mutate(`Location Description`=as.factor(`Location Description`))%>%mutate(District=as.factor(District))%>%mutate(Ward=as.factor(Ward))%>%mutate(`Community Area`=as.factor(`Community Area`))%>%mutate(`FBI Code`=as.factor(`FBI Code`))%>%mutate(Month=as.factor(Month))%>% mutate(Day=as.factor(Day))
```

Let us look at the coorelation among variables 

```{r}
Chicago_NA_Sample <- Chicago_NA[sample(nrow(Chicago_NA),100),]
summary(Chicago_NA_Sample)
```

** Created a smaller dataframe Chicago_NA_Sample with 100 rows only. 

```{r}
ggcorr(Chicago_NA_Sample, label = "TRUE", label_round = 1) 
```

```{r}

```

```{r}
ggplot(Chicago_NA_Sample,aes(Arrest,Beat))+geom_boxplot()
```

```{r}
ggplot(Chicago_NA_Sample, aes(x=Ward, y=Arrest)) + geom_col() 

```

```{r}
ggplot(Chicago_NA_Sample,aes(Arrest,Minute,Colour=Hour))+geom_line()

```


